Self-tunable approximated explicit MPC: Heat exchanger implementation and analysis

### 2024-04-16 paper/JoPC/response_letter_2nd_round.tex

\documentclass[a4paper,10pt]{article}

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{a4wide}
\usepackage{multicol}
\usepackage[ansinew]{inputenc}
\usepackage{color}
\usepackage{url}
%\usepackage{floatflt}
\usepackage{multirow}
\usepackage{rotating, graphicx}

\newcommand{\opt}{^{\star}} % \opt
\newcommand{\tr}{\intercal} % transpose

% RESPONSE LETTER SYNTAX:
% =====================================================
\newcommand{\change}[1]{\textcolor{red}{#1}}

\newcommand{\comment}[1]{
	\begin{itemize}
		\item \textbf{Comment:}\\ \sf{#1}
	\end{itemize}
}

\newcommand{\answer}[1]{
	\begin{itemize}
		\item[] \textbf{Answer:}\\ #1
	\end{itemize}
}

\newcommand{\pointout}[1]{\textcolor{blue}{#1}}
\definecolor{YJ}{rgb}{0.0,0.5,0.0}

% =====================================================

\newcommand{\diam}[1]{\mathrm{diam}\left(#1\right)}

% \newcommand {\mm}[1]{\textcolor{blue}{#1}}

%opening
\title{Response to the review comments}
\author{}
\date{}

\sloppy
\begin{document}

	\maketitle

	We would like to thank the Reviewers for the constructive and thoughtful comments on our manuscript.
	We revised the document based on these comments and we provide point-to-point answers below.
	All changes relative to our initial submission are highlighted in \change{red}. TODO!

	\section*{Answer to Editor}
	We would like to thank the Editor for summarizing substantial and high-quality review comments, which helped us to improve our manuscript's quality. We revised the paper based on the Reviewers' comments and incorporated the suggestions into our manuscript. We reformulated or added the corresponding statements to clarify the ideas.  TODO!

	\section*{Answer to Reviewer 1}
	\comment{ The authors have not addressed some of my comments in the previous submission thoroughly. I have the following follow-up comments:\\

	 The control performance with Qy = 500 is not quantified, and whether the performance benefits of the proposed method reported in Table 2 still remain is not clarified. Since Qy =500 would give a moderately tuned controller, it can provide a better overall performance than the two boundary controllers chosen with "extreme" tuning.
	}

	\answer{		
		We thank the Reviewer for pointing out...

	\begin{table}[h!]
		\begin{center}
			\caption{Control performance criteria.}
			\label{tab:control_performance}
			\begin{tabular}{c|c|c|c|c} 
				Reference step change & $Q_\mathrm{y}$ & SSE [$^{\circ}\mathrm{C}^2$\,s] & $\sigma_{\mathrm{max}}$\,[\%] & $t_{\epsilon}$\,[s]  \\
				\hline
				\multirow{4}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1\,000 & 714 & 33.5 & 16.5 \\
				& 100 & 867 & 16.7 & 12.5 \\ 
				& 500 & \textbf{583} & 17.6 & \textbf{7.5} \\
				& self-tuned & \textbf{678} & \textbf{15.2} & \textbf{9.5}  \\ 
				\hline
				\multirow{4}{*}{ 45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1\,000 & 365 & 47.2 & \textbf{5} \\
				& 100 & 606 & 23.3 & 26.5  \\ 
				& 500 & \textbf{168} & \textbf{18.2} & \textbf{6.5} \\
				& self-tuned & \textbf{248} & \textbf{19.1} & 9.5  \\ 
				\hline
				\multirow{4}{*}{ 50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1\,000 & 245 & \textbf{18.9} & \textbf{6.5}  \\
				& 100 & 398 & 79.6 & 31  \\ 
				& 500 & 211 & \textbf{22.3} & 8 \\
				& self-tuned & \textbf{186} & 24.6 & \textbf{6.5}  \\ 
				\hline
				\multirow{4}{*}{ 45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } & 1\,000 & 1\,024 & 18.4 & 22.5  \\
				& 100 & 1\,402 & 41.9 & 90  \\ 
				& 500 & 1064 & 25.6 & 67 \\
				& self-tuned & \textbf{967} & \textbf{16.5} & \textbf{18.5}   
			\end{tabular}
		\end{center}
	\end{table}

%		\begin{quote}
%			quote			
%		\end{quote}
	}

	\comment{
		For the metrics in Table 2, can the authors also report average values of the three metrics (relative improvements) obtained across the four set point changes considered? The average values would give an idea of the overall control performance. It seems that the overall performance for the settling time would be worse, and only little improvement in the overshoot metric?
	}

	\answer{
		\begin{table}[h!]
			\begin{center}
				\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
				\label{tab:improvement}
				\begin{tabular}{c|c|c|c|c} 
					Reference step change & Comparison subject to the $Q_\mathrm{y}$ setup & $\delta$ SSE\,[\%] & $\delta \sigma_{\mathrm{max}}$\,[\%] & $\delta t_{\epsilon}$\,[\%]  \\
					\hline
					\multirow{2}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 121 &  9 & 74 \\
					& 100 & 28 &  10 & 32 \\  
					\hline
					\multirow{2}{*}{45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1000 & 47 & 147 &$-47$  \\ 
					& 100 & 144 &  22 & 179 \\ 
					\hline
					\multirow{2}{*}{50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 32 &$-23$& 0 \\ 
					& 100 & 114 & 224 & 377 \\
					\hline
					\multirow{2}{*}{45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } &  1000 & 6 & 12 & 22 \\
					& 100 & 45 &  154 & 386 \\
					\hline
					\multirow{2}{*}{Average for all steps} &  1000 & 23 & 64 & 12 \\
					& 100 & 83 &  102 & 243  
				\end{tabular}
			\end{center}
		\end{table}
	}

	\comment{
		All the three performance metric are related to the controlled temperature variable. Why did the authors not consider any metric related to the control input?	
	}

	\comment{
		In addition, the authors only summarize the best case performance improvements in the conclusions. This can be a bit misleading to readers since the proposed method also performs worse (significantly) in some parts of the simulations. So summarizing average values would be more appropriate to give a better picture of the overall performance.
	}

	\comment{
		Based on the responses to my previous comments, it appears that the applicability of the proposed self-tunable explicit MPC method is limited to "small-scale multivariable systems, with decoupled dynamics"? If so, please mention this clearly in the abstract.
	}

	\section*{Answer to Reviewer 2}
	\comment{My comments from the previous round have been fully address. I suggest publication.}

	\answer{		
		We thank the Reviewer again for the review.		
	}

	\bibliographystyle{plain}
	\bibliography{references}

\end{document}

### 2024-04-17 paper/JoPC/response_letter_2nd_round.tex

%opening
\title{Response to the review comments}
\author{}
\date{}

\sloppy
\begin{document}

	\maketitle

	We thank the Reviewers for all comments and suggestions.
	The document was revised according to these comments and we provide point-to-point answers below.
	All changes relative to our initial submission are highlighted in \change{red}.

	\section*{Answer to Editor}
	The authors would like to thank the Editor for collecting the review comments. We revised the paper based on the Reviewers' comments, and incorporated the suggestions
	into our manuscript.

@@

	\section*{Answer to Reviewer 1}
	\comment{ The authors have not addressed some of my comments in the previous submission thoroughly. I have the following follow-up comments:\\

	 The control performance with Qy = 500 is not quantified, and whether the performance benefits of the proposed method reported in Table 2 still remain is not clarified. Since Qy =500 would give a moderately tuned controller, it can provide a better overall performance than the two boundary controllers chosen with "extreme" tuning.
	}

	\answer{		
		We revisited Table~\ref{tab:control_performance} to evaluate the control performance criteria.
		The values highlighted in \textbf{bold} represent the scenarios when the controller with the investigated setup $Q_\mathrm{y} = 500$ performed as the best one, excluding the self-tunable MPC. It can be seen, that the values are the best in 5 out of 12 performance criteria. On the contrary, the self-tuned controller acquired 10 best values out of 12 performance criteria thanks to the tuning based on the varying operating conditions.

		When excluding the boundary controllers and comparing only the self-tunable controller with the setup $Q_\mathrm{y} = 500$, the moderate controller outperforms the self-tunable MPC in half of the performance criteria. It happens mostly in tracking the reference value upwards. Nevertheless, when the reference changes downwards, the self-tuned controller becomes beneficial. Therefore, it supports choosing the tunable control technique.

		\textcolor{blue}{Pozn.: Zvazme teraz dva scenare - bud toto o 500vke zapracujem nejako do clanku, alebo to odbavime nasledovne:}

		To ensure a good readability of the paper, we incline not to branch the case study also in the direction of the moderate setup, i.e., $Q_\mathrm{y} = 500$. We prefer to emphasize the consideration of the self-tuning method based on the revised comment from the first round of review:

		\begin{quote}
			Obviously, if there exists a well-tuned ``universal'' controller that satisfies the requirements on the control performance in the whole range of the considered operating conditions, then the implementation of the self-tuning procedure is out of scope for such control application. Nevertheless, in numerous practical situations, using only one controller with a constant setup leads to poor or just ``satisfactory'' control results, i.e., the reference value is achieved, but with worse control performance, e.g., leading to high overshoots or settling times. When working on our laboratory case study, a set of different setups of penalty matrices was investigated. In every control scenario, the setup was beneficial only in some working conditions (tracking the reference upwards or downwards). Therefore, the closed-loop control performance is improved by introducing the benefits of the self-tuning method based on the two boundary MPC controllers.
		\end{quote}

	\begin{table}[h!]
		\begin{center}
			\caption{Control performance criteria.}
			\label{tab:control_performance}
			\begin{tabular}{c|c|c|c|c} 
				Reference step change & $Q_\mathrm{y}$ & SSE [$^{\circ}\mathrm{C}^2$\,s] & $\sigma_{\mathrm{max}}$\,[\%] & $t_{\epsilon}$\,[s]  \\
				\hline
				\multirow{4}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1\,000 & 714 & 33.5 & 16.5 \\
				& 100 & 867 & 16.7 & 12.5 \\ 
				& 500 & \textbf{583} & 17.6 & \textbf{7.5} \\
				& self-tuned & 678 & 15.2 & 9.5  \\ 
				\hline
				\multirow{4}{*}{ 45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1\,000 & 365 & 47.2 & 5 \\
				& 100 & 606 & 23.3 & 26.5  \\ 
				& 500 & \textbf{168} & \textbf{18.2} & 6.5 \\
				& self-tuned & 248 & 19.1 & 9.5  \\ 
				\hline
				\multirow{4}{*}{ 50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1\,000 & 245 & 18.9 & 6.5  \\
				& 100 & 398 & 79.6 & 31  \\ 
				& 500 & \textbf{211} & 22.3 & 8 \\
				& self-tuned & 186 & 24.6 & 6.5  \\ 
				\hline
				\multirow{4}{*}{ 45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } & 1\,000 & 1\,024 & 18.4 & 22.5  \\
				& 100 & 1\,402 & 41.9 & 90  \\ 
				& 500 & 1064 & 25.6 & 67 \\
				& self-tuned & 967 & 16.5 & 18.5  
			\end{tabular}
		\end{center}
	\end{table}

	}

@@

	\comment{
		For the metrics in Table 2, can the authors also report average values of the three metrics (relative improvements) obtained across the four set point changes considered? The average values would give an idea of the overall control performance. It seems that the overall performance for the settling time would be worse, and only little improvement in the overshoot metric?
	}

	\answer{
		We would like to thank the Reviewer for the comment on the relative improvements. If the values in Table~2 were averaged, it is true that the average settling time would be worse, and the average relative improvement of the overshoot would be very low. However, considering the average of these values would be very strict because of the way how the relative improvements were calculated.

		The values in Table~2 are computed subject to the second best setup for each reference step change individually. The negative numbers represent deterioration compared to the best controller setup in the corresponding reference tracking. It means, that the base subject to which is every value calculated, changes. Therefore, averaging the values of the relative improvements would not be fair, as it would not reflect improvement subject to one controller with a constant setup.

		To avoid the ``unfair'' evaluation of the relative improvement, we extended Table~2 such that it provides the relative improvements subject to each boundary controller individually, see Table~\ref{tab:improvement}.

		\begin{table}[h!]
			\begin{center}
				\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
				\label{tab:improvement}
				\begin{tabular}{c|c|c|c|c} 
					& Comparison subject to the $Q_\mathrm{y}$ setup & $\delta$ SSE\,[\%] & $\delta \sigma_{\mathrm{max}}$\,[\%] & $\delta t_{\epsilon}$\,[\%]  \\
					\hline
					\multirow{2}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 5 &  55 & 42 \\
					& 100 & 22 &  9 & 24 \\  
					\hline
					\multirow{2}{*}{45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1000 & 32 & 59 &$-90$  \\ 
					& 100 & 59 &  18 & 64 \\ 
					\hline
					\multirow{2}{*}{50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 24 &$-30$& 0 \\ 
					& 100 & 53 & 69 & 79 \\
					\hline
					\multirow{2}{*}{45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } &  1000 & 6 & 11 & 18 \\
					& 100 & 31 &  61 & 79 \\
					\hline
					\multirow{2}{*}{Average across all steps} &  1000 & 17 & 24 & -7 \\
					& 100 & 41 &  39 & 62  
				\end{tabular}
			\end{center}
		\end{table}
	}

@@

	\comment{
		In addition, the authors only summarize the best case performance improvements in the conclusions. This can be a bit misleading to readers since the proposed method also performs worse (significantly) in some parts of the simulations. So summarizing average values would be more appropriate to give a better picture of the overall performance.
	}

	\comment{
		Based on the responses to my previous comments, it appears that the applicability of the proposed self-tunable explicit MPC method is limited to "small-scale multivariable systems, with decoupled dynamics"? If so, please mention this clearly in the abstract.
	}

	\section*{Answer to Reviewer 2}
	\comment{My comments from the previous round have been fully address. I suggest publication.}

	\answer{		
		We thank the Reviewer for reviewing our revised manuscript.		
	}

### 2024-04-17 paper/JoPC/response_letter_2nd_round.tex

	\answer{
		We would like to thank the Reviewer for the comment on the relative improvements. If the values in Table~2 were averaged, it is true that the average settling time would be worse, and the average relative improvement of the overshoot would be very low. However, considering the average of these values would be very strict because of the way how the relative improvements were calculated.

		The values in Table~2 are computed subject to the second best setup for each reference step change individually. The negative numbers represent deterioration compared to the best controller setup in the corresponding reference tracking. It means, that the base subject to which is every value calculated, changes. Therefore, averaging the values of the relative improvements would not be fair, as it would not reflect improvement subject to one controller with a constant setup.

		To avoid the ``unfair'' evaluation of the relative improvement, we extended Table~2 such that it provides the relative improvements subject to each boundary controller individually, see Table~\ref{tab:improvement}.

		\begin{table}[h!]
			\begin{center}
				\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
				\label{tab:improvement}
				\begin{tabular}{c|c|c|c|c} 
					& Comparison subject to the $Q_\mathrm{y}$ setup & $\delta$ SSE\,[\%] & $\delta \sigma_{\mathrm{max}}$\,[\%] & $\delta t_{\epsilon}$\,[\%]  \\
					\hline
					\multirow{2}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 5 &  55 & 42 \\
					& 100 & 22 &  9 & 24 \\  
					\hline
					\multirow{2}{*}{45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1000 & 32 & 59 &$-90$  \\ 
					& 100 & 59 &  18 & 64 \\ 
					\hline
					\multirow{2}{*}{50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 24 &$-30$& 0 \\ 
					& 100 & 53 & 69 & 79 \\
					\hline
					\multirow{2}{*}{45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } &  1000 & 6 & 11 & 18 \\
					& 100 & 31 &  61 & 79 \\
					\hline
					\multirow{2}{*}{Average across all steps} &  1000 & 17 & 24 & -7 \\
					& 100 & 41 &  39 & 62  
				\end{tabular}
			\end{center}
		\end{table}

			\begin{table}[h!]
		\begin{center}
			\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
			\label{tab:improvement2}
			\begin{tabular}{c|c|c|c|c} 
				& Comparison subject to the $Q_\mathrm{y}$ setup & $\delta$ SSE\,[\%] & $\delta \sigma_{\mathrm{max}}$\,[\%] & $\delta t_{\epsilon}$\,[\%]  \\
				\hline
				\multirow{3}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 5 &  55 & 42 \\
				& 100 & 22 &  9 & 24 \\ 
				& 500 & -16 & 13 & -27 \\
				\hline
				\multirow{3}{*}{45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1000 & 32 & 59 &$-90$  \\ 
				& 100 & 59 &  18 & 64 \\ 
				& 500 & -48 & -5 & -46 \\
				\hline
				\multirow{3}{*}{50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 24 &$-30$& 0 \\ 
				& 100 & 53 & 69 & 79 \\
				& 500 & 12 & -10 & 19 \\
				\hline
				\multirow{3}{*}{45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } &  1000 & 6 & 11 & 18 \\
				& 100 & 31 &  61 & 79 \\
				& 500 & 9 & 36 & 72 \\
				\hline
				\multirow{3}{*}{Average across all steps} &  1000 & 17 & 24 & -7 \\
				& 100 & 41 &  39 & 62 \\ 
				& 500 & -11 & 8 &  5
			\end{tabular}
		\end{center}
	\end{table}

@@

	}

	\comment{
		All the three performance metric are related to the controlled temperature variable. Why did the authors not consider any metric related to the control input?	
	}

	\answer{
		We thank the Reviewer for commenting on the control performance criteria related to the control input. Although we considered such an evaluation, we chose not to include the criteria for the following reasons. This kind of criteria usually reflects some costs or energy demands associated with the control.

		The first kind of criterium related to the control input could be intuitively some (potentially weighted) ISU (integral-of-squared input) or, in a discrete-time manner, SSU (sum-of-squared input). These criteria correspond to the summing of the control input, i.e., voltage, to set the operating speed of the pump. The operating speed of the pump itself does not play any significant role in the costs of the control. Therefore, it could be converted to the amount of liquid transported to the heat exchanger during the whole control experiment. Nevertheless, the heating medium is dosed to the heat exchanger plant as a recycle, i.e., the same amount of heating medium is re-used during operation, which would represent fictitious performance criteria not related to the actual plant setup.

		Another criterion related to the control input we considered was the energy consumed to pre-heat the heating medium in the heater. However, this is not directly associated with the MPC's control inputs. The consumed energy corresponds to the control involving the proportional controller, which sets the electric power. Therefore, we decided not to include such a criterion as well.
	}

@@

	\comment{
		In addition, the authors only summarize the best case performance improvements in the conclusions. This can be a bit misleading to readers since the proposed method also performs worse (significantly) in some parts of the simulations. So summarizing average values would be more appropriate to give a better picture of the overall performance.
	}

	\answer{
		We thank the Reviewer for pointing out the misleading communication of the control results...
	}

### 2024-04-21 paper/JoPC/main_3rd_round_marked.tex
reponse letter 2 - bugfix, main_3rdround.tex created

		\begin{abstract}
			%%max 250 words!			
			The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any interventions of the control engineer during operation in order to \change{retune the controller subject to the changed working conditions.} Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. This extension provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetric behavior of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning procedure. The designed self-tunable controller improved control performance by decreasing sum-of-squared control error, maximal overshoots/undershoots, and settling time compared to the conventional control strategy based on a single (non-tunable) controller. 
		\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%Research highlights
%		\begin{highlights}
%			 %3-5 highlights, every one should not exceed 85 characters including spaces
%			\item Self-tunable control technique does not need any intervention during control
%			\item Tuning based on operating conditions compensates for the nonlinear plant behavior
%			\item Implementation of the proposed tuning method on a laboratory heat exchanger 
%			\item Control performance improved compared to utilizing explicit MPC with constant setup
%		\end{highlights}

@@

As can be seen in Table~\ref{tab:control_performance}, the real-time self-tuning of the explicit MPC controller helped to improve two to three criteria when tracking each reference value. 	
The relative improvement in the percentage, denoted by $\delta$, of using the self-tunable controller computed subject to the second best setup is summarized in Table~\ref{tab:improvement} for each reference step change separately. The negative numbers represent deterioration compared to the best controller setup in the corresponding reference tracking.

\begin{table}[h!]
	\begin{center}
		\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
		\label{tab:improvement}
		\begin{tabular}{c|c|c|c} 
			Reference step change & $\delta$ SSE\,[\%] & $\delta \sigma_{\mathrm{max}}$\,[\%] & $\delta t_{\epsilon}$\,[\%]  \\
			\hline
			35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C &  5 &  9 & 24 \\ 
			% \hline
			45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 32 & 18 &$-90$  \\ 
			% \hline
			50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C & 24 &$-30$& 0 \\ 
			% \hline
			45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C &  6 & 11 & 18   
		\end{tabular}
	\end{center}
\end{table}
%\begin{table}[h!]
%\begin{center}
%\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
%\label{tab:improvement}
%\begin{tabular}{c|c|c|c|c} 
%Reference step change & $\Delta ISE$\,[\%] & $\Delta \sigma_{\mathrm{max}}$\,[\%] & $\Delta t_{\epsilon}$\,[\%] & $\Delta V$\,[\%] \\
%\hline
%35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C & 5.04 & 8.77 & 24.00 & -12.26 \\ 
%\hline
%45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 32.05 & 17.72 & -90.00 & -34.54 \\ 
%\hline
%45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 24.08 & -29.97 & 0 & -4.48 \\ 
%\hline
%45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C & 5.57 & 10.53 & 17.78 & -18.28  
%\end{tabular}
%\end{center}
%\end{table}

Compared to the considered non-self-tunable controllers, the control trajectories and the evaluated quality criteria confirmed the improved control performance for the reference tracking control problem of the heat exchanger with the non-linear and asymmetric behavior. Implementing a self-tunable explicit MPC controller leads to improved control performance in the most analyzed quality criteria, see Table~\ref{tab:improvement}.	 
The squared-error-based criterion (SSE) improved in every control scenario within a range $[5-32]$\,\%, the maximal overshoots/undershoots $\sigma_{\mathrm{max}}$ reduced up to 18\,\%, and the settling time $t_{\epsilon}$ decreased up to 24\,\%.

@@

	\change{Obviously, if there exists a well-tuned ``universal'' controller that satisfies the requirements on the control performance in the whole range of the considered operating conditions, then the implementation of the self-tuning procedure is out of scope for such control application. Nevertheless, in numerous practical situations, using only one controller with a constant setup leads to poor or just ``satisfactory'' control results, i.e., the reference value is achieved, but with worse control performance, e.g., leading to high overshoots or settling times. When working on our laboratory case study, a set of different setups of penalty matrices was investigated. In every control scenario, the setup was beneficial only in some working conditions (tracking the reference upwards or downwards). Therefore, the closed-loop control performance is improved by introducing the benefits of the self-tuning method based on the two boundary MPC controllers.}

@@

	To properly investigate the control results, the control performance was also judged quantitatively using a set of quality criteria. The self-tunable control approach outperformed the conventional control strategy handling just a single controller, i.e., non-tunable controller, by minimizing the sum-of-squared control error up to 32\,\%, reducing maximal overshoots/undershoots up to 18\,\%, and decreasing settling time up to 24\,\%.

>>>>

		\begin{abstract}
			%%max 250 words!			
			The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any interventions of the control engineer during operation in order to \change{retune the controller subject to the changed working conditions.} Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. This extension provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. \textcolor{red}{The self-tunable technique is suitable for SISO (single-input and single-output) systems and is straighforward to implement for MIMO (multiple-input and multiple-output) systems with decoupled dynamics.} The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetric behavior of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning procedure. The designed self-tunable controller improved control performance by decreasing sum-of-squared control error, maximal overshoots/undershoots, and settling time compared to the conventional control strategy based on a single (non-tunable) controller. 
		\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%Research highlights
%		\begin{highlights}
%			 %3-5 highlights, every one should not exceed 85 characters including spaces
%			\item Self-tunable control technique does not need any intervention during control
%			\item Tuning based on operating conditions compensates for the nonlinear plant behavior
%			\item Implementation of the proposed tuning method on a laboratory heat exchanger 
%			\item Control performance improved compared to utilizing explicit MPC with constant setup
%		\end{highlights}

@@

As can be seen in Table~\ref{tab:control_performance}, the real-time self-tuning of the explicit MPC controller helped to improve two to three criteria when tracking each reference value. 	
\textcolor{red}{The relative improvement in the percentage, denoted by $\delta$, of using the self-tunable controller is summarized in Table~\ref{tab:improvement} for each reference step change separately. The values were computed as the difference between two criteria values corresponding to the optimal and self-tunable MPC, referred to the self-tunable MPC. The negative numbers represent deterioration of the specific performance criterion in the corresponding reference tracking.}

\begin{table}[h!]
	\begin{center}
		\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
		\label{tab:improvement}
		\begin{tabular}{c|c|c|c|c} 
			& Comparison with $Q_\mathrm{y}$ setup & $\delta$ SSE\,[\%] & $\delta \sigma_{\mathrm{max}}$\,[\%] & $\delta t_{\epsilon}$\,[\%]  \\
			\hline
			\multirow{3}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 5 &  121 & 74 \\
			& 100 & 28 &  10 & 32 \\ 
			\hline
			\multirow{3}{*}{45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1000 & 47 & 147 &$-47$  \\ 
			& 100 & 144 &  22 & 179 \\
			\hline
			\multirow{3}{*}{50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 32 &$-23$& 0 \\ 
			& 100 & 114 & 224 & 377 \\
			\hline
			\multirow{3}{*}{45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } &  1000 & 6 & 12 & 22 \\
			& 100 & 45 &  154 & 386 \\
			\hline
			\multirow{3}{*}{Average} &  1000 & 23 & 64 & 12 \\
			& 100 & 83 &  102 & 244
		\end{tabular}
	\end{center}
\end{table}	
%\begin{table}[h!]
%\begin{center}
%\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
%\label{tab:improvement}
%\begin{tabular}{c|c|c|c|c} 
%Reference step change & $\Delta ISE$\,[\%] & $\Delta \sigma_{\mathrm{max}}$\,[\%] & $\Delta t_{\epsilon}$\,[\%] & $\Delta V$\,[\%] \\
%\hline
%35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C & 5.04 & 8.77 & 24.00 & -12.26 \\ 
%\hline
%45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 32.05 & 17.72 & -90.00 & -34.54 \\ 
%\hline
%45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 24.08 & -29.97 & 0 & -4.48 \\ 
%\hline
%45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C & 5.57 & 10.53 & 17.78 & -18.28  
%\end{tabular}
%\end{center}
%\end{table}

Compared to the considered non-self-tunable controllers, the control trajectories and the evaluated quality criteria confirmed the improved control performance for the reference tracking control problem of the heat exchanger with the non-linear and asymmetric behavior. Implementing a self-tunable explicit MPC controller leads to improved control performance in the most analyzed quality criteria, see Table~\ref{tab:improvement}.	 
\textcolor{red}{In average, the control performance criteria improved compared to the upper and lower boundary MPC respectively as follows: the squared-error-based criterion (SSE) reduced by 23\% and 83\%, the maximal overshoot/undershoot $\sigma_{\mathrm{max}}$ reduced by 64\% and 102\%, and the settling time $t_{\epsilon}$ reduced by 12\% and 244\%.}

@@

	\textcolor{red}{Obviously, if there exists a well-tuned ``universal'' controller that satisfies the requirements on the control performance in the whole range of the considered operating conditions, then the implementation of the self-tuning procedure is out of scope for such control application. Nevertheless, in numerous practical situations, using only one controller with a constant setup leads to poor or just ``satisfactory'' control results, i.e., the reference value is achieved, but with worse control performance, e.g., leading to high overshoots or settling times. When working on our laboratory case study, a set of different setups of penalty matrices was investigated. In each control scenario, the setup was beneficial only in some working conditions (tracking the reference upwards or downwards). Therefore, the closed-loop control performance is significantly improved by introducing the benefits of the self-tuning method based on the two well-tuned boundary MPC controllers.}

@@

	To properly investigate the control results, the control performance was also judged quantitatively using a set of quality criteria. The self-tunable control approach outperformed the conventional control strategy handling just a single controller, i.e., non-tunable controller. \textcolor{red}{In average, the control performance criteria improved compared to the upper and lower boundary MPC respectively as follows: the squared-error-based criterion (SSE) reduced by 23\,\% and 83\,\%, the maximal overshoot/undershoot $\sigma_{\mathrm{max}}$ reduced by 64\,\% and 102\,\%, and the settling time $t_{\epsilon}$ reduced by 12\,\% and 244\,\%.}
